Exploring the Motives and Driving Forces Behind Modifications to APIs From a Software Evolution Perspective

### 2015-05-04 thesis/paper.tex
Removed Gothenburg, Sweden from auth block
Added missing fullstops in discussion

\author{
% 1st. author
\alignauthor
John Burchell\\
       \affaddr{Computer Science and Engineering}\\
       \affaddr{University of Gothenburg}\\
       \affaddr{Gothenburg, Sweden}\\
       \email{john.a.burchell@gmail.com}
% 2nd. author
\alignauthor
William Granli\\
       \affaddr{Computer Science and Engineering}\\
       \affaddr{University of Gothenburg}\\
       \affaddr{Gothenburg, Sweden}\\
       \email{william.granli@gmail.com}
\and
% 3rd. author
\alignauthor Imed Hammouda\titlenote{Co-author}\\
       \affaddr{Computer Science and Engineering}\\
       \affaddr{University of Gothenburg}\\
       \affaddr{Gothenburg, Sweden}\\
       \email{imed.hammouda@gu.se}
% 4th 
\alignauthor Eric Knauss\titlenote{Co-author}\\
       \affaddr{Computer Science and Engineering}\\
       \affaddr{University of Gothenburg}\\
       \affaddr{Gothenburg, Sweden}\\
       \email{imed.hammouda@gu.se}
}

@@

\noindent
\textbf{Adding functionality is a prime driver of API evolution  }
Overwhelmingly, the primary driving factor behind API evolution was the desire for new API functionality. We found strong evidence to suggest that most, if not all of the additions, re-implementations and functional changes were undertaken to include additional functionality requested both internally and by the API users. This focus upon functionality was also emphasised in part due to a business goal. The more features the API can provide, the more applications developers can create for the cameras. This has a knock-on effect of increasing the appeal and therefore value of the products.

%API users should be included in the design process?
\smallskip \noindent
\textbf{Maintainability does not drive change  } 
While maintainability is important, it is not a primary driving factor for change. We found that where maintainability changed, functionality was also altered. For example, we predicted that the main intents to be enhancive which was correct. Furthermore, of the changes identified, none were considered to be primarily groomative, groomative changes were only identified as sub-category changes. Changes to the example code is further evidence of this. The example code was updated purely to reflect the changes in functionality and not as a groomative effort.

\smallskip \noindent
\textbf{Aiding development of applications  } 
Aiding users of the API to create correct and robust applications is a very important evolution factor for Axis. Changes between the versions brought about improvements to error handling, the example code and the way in which the API utilises interfaces. This, coupled with removal of unwanted functionality and unsafe operations, has allowed the API to grow and become much more user friendly. As mentioned previously, this could increase the amount of applications being created for the camera and thus could increase its value.

\smallskip \noindent
\textbf{Removing functionality  }
Previous literature has stated that deprecations should be favoured over removals from API for many reasons except in specific circumstances. The successful removal of the burst function seems to indicate that this advice is well worth following. Indeed, axis only removed this function from the API after an internal review found that the cost of removing it would outweigh potential drawbacks to keeping it in the API. It is therefore recommended to only remove something if the effects of removing it have been well investigated and that the removal is rather small in scale.

\smallskip \noindent
\textbf{Updative changes do not drive change  }
Changes to documentation and example code are important, especially if the API is drastically changing between versions. However, these updative changes are drivers themselves, rather, they make up part of a larger change as a whole. 
%^Probably needs more to it?

\smallskip \noindent
\textbf{Hardware was not a driving factor  }
Interestingly, we found that hardware was not a direct driver of change. Initially we had thought that it would be one of the strongest drivers of change, for example, new versions of hardware become available or new hardware is added to the camera requiring something in the API to access it. This does seem to suggest that the hardware for the camera is sufficiently stable such that it does not require changes and that those changes do not need to be reflected in new API versions.

>>>>

\author{
% 1st. author
\alignauthor
John Burchell\\
       \affaddr{Computer Science and Engineering}\\
       \affaddr{University of Gothenburg}\\
       \email{john.a.burchell@gmail.com}
% 2nd. author
\alignauthor
William Granli\\
       \affaddr{Computer Science and Engineering}\\
       \affaddr{University of Gothenburg}\\
       \email{william.granli@gmail.com}
\and
% 3rd. author
\alignauthor Imed Hammouda\titlenote{Co-author}\\
       \affaddr{Computer Science and Engineering}\\
       \affaddr{University of Gothenburg}\\
       \email{imed.hammouda@gu.se}
% 4th 
\alignauthor Eric Knauss\titlenote{Co-author}\\
       \affaddr{Computer Science and Engineering}\\
       \affaddr{University of Gothenburg}\\
       \email{imed.hammouda@gu.se}
}

@@

\noindent
\textbf{Adding functionality is a prime driver of API evolution.  }
Overwhelmingly, the primary driving factor behind API evolution was the desire for new API functionality. We found strong evidence to suggest that most, if not all of the additions, re-implementations and functional changes were undertaken to include additional functionality requested both internally and by the API users. This focus upon functionality was also emphasised in part due to a business goal. The more features the API can provide, the more applications developers can create for the cameras. This has a knock-on effect of increasing the appeal and therefore value of the products.

%API users should be included in the design process?
\smallskip \noindent
\textbf{Maintainability does not drive change.  } 
While maintainability is important, it is not a primary driving factor for change. We found that where maintainability changed, functionality was also altered. For example, we predicted that the main intents to be enhancive which was correct. Furthermore, of the changes identified, none were considered to be primarily groomative, groomative changes were only identified as sub-category changes. Changes to the example code is further evidence of this. The example code was updated purely to reflect the changes in functionality and not as a groomative effort.

\smallskip \noindent
\textbf{Aiding development of applications.  } 
Aiding users of the API to create correct and robust applications is a very important evolution factor for Axis. Changes between the versions brought about improvements to error handling, the example code and the way in which the API utilises interfaces. This, coupled with removal of unwanted functionality and unsafe operations, has allowed the API to grow and become much more user friendly. As mentioned previously, this could increase the amount of applications being created for the camera and thus could increase its value.

\smallskip \noindent
\textbf{Removing functionality.  }
Previous literature has stated that deprecations should be favoured over removals from API for many reasons except in specific circumstances. The successful removal of the burst function seems to indicate that this advice is well worth following. Indeed, axis only removed this function from the API after an internal review found that the cost of removing it would outweigh potential drawbacks to keeping it in the API. It is therefore recommended to only remove something if the effects of removing it have been well investigated and that the removal is rather small in scale.

\smallskip \noindent
\textbf{Updative changes do not drive change.  }
Changes to documentation and example code are important, especially if the API is drastically changing between versions. However, these updative changes are drivers themselves, rather, they make up part of a larger change as a whole. 
%^Probably needs more to it?

\smallskip \noindent
\textbf{Hardware was not a driving factor.  }
Interestingly, we found that hardware was not a direct driver of change. Initially we had thought that it would be one of the strongest drivers of change, for example, new versions of hardware become available or new hardware is added to the camera requiring something in the API to access it. This does seem to suggest that the hardware for the camera is sufficiently stable such that it does not require changes and that those changes do not need to be reflected in new API versions.
